Overlapping Batches for the Assessment of Solution Quality in Stochastic Programs

### 2014-07-03 Referee2_Response.tex

%%%%%%%%%%%%%%%%%
%% Need to be careful with the minor changes. Will these be affected?
%% Need to CHECK AT THE VERY END... FOR ANY CHANGES on TOP OF these...
%%    - GB
%%%%%%%%%%%%%%%%%

\noi  
{\large \bf General Comments:}
\medskip

\begin{itemize}
\item[1.] \textit{The main contributions of the paper need to be stated more clearly and succinctly.}
\end{itemize}

\noindent 
xxx  We agree. 
We changed the introduction to clearly state the contributions of the paper and \ldots  and the presentation of the paper significantly improved.  Thank you.
\medskip

@@

\begin{itemize}
\item[2.] \textit{It is not necessary to go into such detail in the OBM review.}
\end{itemize}

\smallskip

\noindent 
To shorten and streamline the OBM review, and to introduce the main results used in the paper, we combined the previous two sections---\S 2.1 (Overlapping Batch Means) and \S 3.1 (Variably Overlapping Batches)---with other explanations on OBM found later in the paper. 
The combined review appears in the background section, \S 2.1 (Batch Means). 
To improve readability, we decided to create two subsections: \S 2.1.1 (Nonoverlapping Batch Means) and \S 2.1.2 (Overlapping Batch Means). 
While we removed a number of equations and shortened the text, we kept some explanations that are necessary for a reader who is not familiar with OBM (e.g., readers with an optimization focus) to follow the main ideas. 
These basic explanations appear, for instance, in the first paragraph of \S 2.1.2 (Overlapping Batch Means).  
We hope to be able to reach a broader audience this way. 
As a result of these changes, 
\begin{itemize}
	\item[(i)]  instead of former equations (2) and (5), both on overlapping batch means variance estimator, we have only one equation (2);

	\item[(ii)]  the confidence interval estimation using OBM is discussed only once;

	\item[(iii)] the former equation (3) is removed;

	\item[(iv)] the discussion that appeared at the beginning of \S 4.4 (previously: Variance Reduction, now: Asymptotically Lower Variances) is moved to this review section. 
\end{itemize}

\noi
With these changes, the OBM review became more streamlined, focusing on the main results used later in the paper. 
Thank you.
\bigskip

@@

\begin{itemize}
\item[] \textit{4, 51: ``discuss overlapping''}
\end{itemize}

\noi
Done.  
\medskip

\begin{itemize}
\item[] \textit{4, 54 (and elsewhere): When you refer to ``variable amounts of overlap'', that is really nothing new. Certainly, any overlapping estimator will have end effects in which the overlap amounts differ from those in the ``middle'' region of the data.}
\end{itemize}

\noi
%You are right.
We meant to say that the degree of overlap is parametrized and this parameter can be chosen by the user. 
We reworded this sentence to reflect this. 
We also went through the paper, removed other occurrences, and avoided using this term in the new text added. 
\medskip

@@

\begin{itemize}
\item[] \textit{4, 57 (and elsewhere): ``classical case'' refers to nonoverlapping batch means, right?}
\end{itemize}

\noi
No; by `classical case', we meant the use of overlapping batches in simulation output analysis (as opposed to the assessment of solution quality in stochastic programming). 
We reworded this sentence to make it clear. 
Please note that based on your earlier comments, we reworded this sentence to avoid using the term `variance reduction'. 
We also went through the paper and removed several uses of the word `classical' that referred to nonoverlapping batches to reduce ambiguity.
%(as these were referring to nonoverlapping batches). 
%In the current version of the paper, `classical case' is used to only indicate  the use of overlapping batches in simulation output analysis. 
\medskip

\begin{itemize}
\item[] \textit{5, 12: Does $\sigma^2$ refer to the marginal variance or the ``variance parameter'' (sum of the covariances at all lags)?}
\end{itemize}

\noi
Variance parameter. 
We revised this sentence and added a footnote defining the variance parameter for readers who might not be familiar with this term (e.g., readers with an optimization focus). 
\medskip

@@

\begin{itemize}
\item[] \textit{5, 43: You could have introduced the notation ``CI'' earlier.}
\end{itemize}

\noi
You are right. 
We now introduce this abbreviation the first time we use the term `confidence interval' in the introduction. 
From this point on, we use the abbreviated form `CI' throughout the paper, except for the title of Section 4.5 (Asymptotic Validity of the Confidence Intervals). 
\medskip

\begin{itemize}
\item[] \textit{5: You can significantly reduce the discussion on OBM here and in \S 3.1. 
Just let the reader know the main results that you'll use.}
\end{itemize}

\noi
We removed \S 3.1 (Variably Overlapping Batches) from the paper. 
We merged this section with \S 2.1.2 (Overlapping Batch Means), focusing on the results used in the paper. 
We believe the presentation of the paper improved as a result. 
Thank you for this suggestion. 
\medskip

@@

\begin{itemize}
\item[] \textit{6, 29: Note that you use ``Note that'' too often.}
\end{itemize}

\noi
We went through the paper and removed most of the ``note that'' phrases. 
We decided to leave a couple as is (mainly as ``notice that'') because they highlight important characteristics.  
\medskip

\begin{itemize}
\item[] \textit{6, 30: Explicitly say what the d.f.\ in (2) is.}
\end{itemize}

\noi
We changed this sentence and directly referred to the denominator term.
We decided to avoid the degrees of freedom here. 
This is because we wanted to focus on the degrees of freedom for confidence interval construction, which is discussed in both the theoretical and computational results.\smallskip

\noi 
{\it Note:} This section has been restructured based on your comments. 
Equation (2) now gives the general overlapping variance estimator with the parametrized amount of overlap. 
\medskip

@@

\begin{itemize}
\item[] \textit{7, 46: Does $\gamma$ need to be an integer? 
You seem to be using it that way, so maybe state the requirement.}
\end{itemize}

\noi
Yes; we explicitly stated this in a sentence after we introduce $\gamma$.\smallskip

\noi 
{\it Note:} This sentence now appears in Section 2.1.2 (Overlapping Batch Means). 
\medskip

@@

\begin{itemize}
\item[] \textit{8, 4 (and elsewhere): Define notations such as floor and ceiling functions.}
\end{itemize}

\noi
Done. 
%We introduced the floor function before its first use, and we introduced the ceiling function right after its first use.
\smallskip

\noi 
{\it Note:} Both definitions appear in Section 2.1.2 (Overlapping Batch Means).
\medskip

@@

\begin{itemize}
\item[] \textit{8, 11: In addition to the fact that I don't think ``variance reduction'' is quite the right term, the phrase ``variance reduction in this estimator'' is awkward.}
\end{itemize}

\noi
Reworded.\smallskip

\noi 
{\it Note:} This sentence now appears in Section 2.1.2 (Overlapping Batch Means). 
\medskip

@@

\begin{itemize}
\item[] \textit{8, 53--57: The various OBM references should have been discussed earlier.}
\end{itemize}

\noi
You are right. 
We moved these references---in particular, Damerdji (1994, 1995)---to Section 2.1.2 (Overlapping Batch Means) and added a brief discussion.  
\medskip

@@

\noi
Instead of adding another step before the first inequality in the proof of Lemma 1, we added explanations at the beginning of the proof to make this inequality clear.   
\medskip

\begin{itemize}
\item[] \textit{12, 52: ``$n_l VD_l$ has been studied''}
\end{itemize}

\noi
Done.
\medskip

\begin{itemize}
\item[] \textit{13, 12 (and elsewhere): I may have missed this, but has the notation $\sigma_{\hat{x}}$ been established yet?}
\end{itemize}

\noi
xxx 
This notation was defined before Theorem 1. 
We moved the definition of $\sigma_{\hat{x}}$ and $\mu_{\hat{x}}$ to \ldots to improve clarity. 
\medskip

\begin{itemize}
\item[] \textit{In any case, some of the verbiage in this paragraph is a bit repetitive.}
\end{itemize}

\noi
We moved the discussion at the first half of this paragraph to the new, combined OBM review in \S 2.1.2 (Overlapping Batch Means). 
We referred to equation (3) here to remind the readers the basis of the proof.
The second half of this paragraph deals with the batching structure, and this discussion is moved to the paragraph after the theorem's proof. 
\medskip

### 2014-07-03 omrp_lb_rev.tex
More Ref2 responses

Our aim is to apply the idea of overlapping batches to MRP.  
Overlapping could help to obtain better variance estimators with simple data reuse, resulting in improved CIs.
%Especially for problems where sampling error is dominant, this simple data reuse method could yield better estimators.
We note several differences in this setting compared to the classical simulation setting.  
In simulation output analysis, the point of interest is that of estimating the variance of the sample mean of a covariance stationary process.  
We are interested in estimating the variance of an optimality gap estimator.  
An optimality gap estimator not only has a sample mean $(\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti))$ but also a minimized sample mean $(\zs_m)$.  
Minimization changes the statistical properties of sample means, complicating the asymptotic analysis.  
%For instance, the central limit theorem may not hold for a minimized sample mean, even though it holds for each $\x \in \X$.  
%Therefore, analysis of the asymptotic benefits of overlapping becomes more complicated.
We overcome this technical difficulty by approximating the optimality gap estimators by their nonoptimized counterparts (see \S \ref{sec:theory}).  
The nonoptimized counterparts are (regular) overlapping batch means estimators (without any optimization present); therefore, they have the desired statistical properties of approximately the same bias and lower variances compared to their nonoverlapping versions. 
We establish convergence of the optimality gap estimators that use overlapping to their nonoptimized counterparts. 
Another difference between our setting and simulation output analysis is that once the data are generated through a simulation, they can be reused without much additional computational effort to obtain the OBM variance estimator.  
In this setting, due to the need to solve a sampling problem (SP$_m$), the computational effort may increase with data reuse.  
This can be alleviated in two ways. 
First, near-optimal variance reduction can be obtained by partially overlapping the batches \citep{Welch1987,Song1992}.  
Partial overlapping results in a smaller number of batches; hence, fewer optimization problems need to be solved.
Second, in many solution methods, warm-starting can be used to quickly solve the sampling approximations with overlapping samples, considerably reducing solution time.
In this paper, we consider variable amounts of overlap and examine this both theoretically and empirically. 
% THIS LAST SENTENCE NEEDS TO CHANGE!

@@

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$, relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right).
\end{equation}
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/4$), for instance, the variance is 75\% of the nonoverlapping counterpart.

@@

The optimality gap estimator for each batch $\gb_{lj}$ is defined like (\ref{eq:mrp_gap}) for general values of the nonoverlap parameter $\gamma_l$.  
We just removed the minimization in (\ref{eq:mrp_gap}) and used directly an optimal solution of $\xs_j$ of batch $j$.  
The overall mean, $\gbb_l$, is defined a little differently.  
Here, $\gbb_l$ still uses each observation $i = 1, 2, \dots, n_l$ but also makes use of {\it all} the information collected throughout the batches.  
That is, if observation $\xiti$ is used in $|B(i)|$ number of batches, all the optimal solutions $\xs_j$ corresponding to each batch $j \in B(i)$ are used for the lower bound estimator.  
Then, $VG_l$ is defined in a similar fashion for the variably overlapping batches variance estimator (\ref{eq:variance_gamma}).

@@

The internal optimization in the batches in (\ref{eq:gbar}) makes a straightforward statistical analysis of the behavior of the estimators difficult.  
To overcome this problem, we introduce the following unbiased optimality gap estimators:
\begin{align}
	\db_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs,\xit^{\gamma_l(j-1)+i}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:dbar} \\
	\dbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \left[ f(\xh,\xiti) - f(\xs,\xiti) \right], \label{eq:dbb} \\
	VD_l & = \frac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\db_{lj} - \dbb_l)^2. \label{eq:vd}
\end{align}
These are essentially the same as the variably overlapping batches estimators in \S \ref{ssec:overlap} with $y^i = f(\xh,\xiti) - f(\xs,\xiti)$.  
These are also defined identically to the original estimators (\ref{eq:gbar})--(\ref{eq:vg}) with the exception that $\xs_j$ from (\ref{eq:gbar}) and (\ref{eq:gbb}) is replaced by an optimal solution $\xs$ in (\ref{eq:dbar}) and (\ref{eq:dbb}).  
With $\xh$ and $\xs$ fixed, estimators (\ref{eq:dbar})--(\ref{eq:vd}) have the same statistical properties as variably overlapping batches estimators in \S \ref{ssec:overlap}.  
The optimal solution $\xs$ is not known; however, the estimators (\ref{eq:dbar})--(\ref{eq:vd}) are used only to show convergence of (\ref{eq:gbar})--(\ref{eq:vg}) and not necessary for carrying out the OMRP algorithm.  
Throughout the rest of this section, we assume that the optimality gap estimators (\ref{eq:gbar})--(\ref{eq:vg}) and their nonoptimized counterparts (\ref{eq:dbar})--(\ref{eq:vd}) use identical samples and values of $n_
l$, $m_l$, and $\gamma_l$.  \smallskip

>>>>

Our aim is to apply the idea of overlapping batches to MRP.  
Overlapping could help to obtain better variance estimators with simple data reuse, resulting in improved CIs.
%Especially for problems where sampling error is dominant, this simple data reuse method could yield better estimators.
We note several differences in this setting compared to the classical simulation setting.  
In simulation output analysis, the point of interest is that of estimating the variance of the sample mean of a covariance stationary process.  
We are interested in estimating the variance of an optimality gap estimator.  
An optimality gap estimator not only has a sample mean $(\frac{1}{m} \sum_{i=1}^m f(\xh,\xiti))$ but also a minimized sample mean $(\zs_m)$.  
Minimization changes the statistical properties of sample means, complicating the asymptotic analysis.  
%For instance, the central limit theorem may not hold for a minimized sample mean, even though it holds for each $\x \in \X$.  
%Therefore, analysis of the asymptotic benefits of overlapping becomes more complicated.
We overcome this technical difficulty by approximating the optimality gap estimators by their nonoptimized counterparts (see \S \ref{sec:theory}).  
The nonoptimized counterparts are (regular) overlapping batch means estimators (without any optimization present); therefore, they have the desired statistical properties of approximately the same bias and lower variances compared to their nonoverlapping versions. 
We establish convergence of the optimality gap estimators that use overlapping to their nonoptimized counterparts. 
Another difference between our setting and simulation output analysis is that once the data are generated through a simulation, they can be reused without much additional computational effort to obtain the OBM variance estimator.  
In this setting, due to the need to solve a sampling problem (SP$_m$), the computational effort may increase with data reuse.  
This can be alleviated in two ways. 
First, near-optimal variance reduction can be obtained by partially overlapping the batches \citep{Welch1987,Song1992}.  
Partial overlapping results in a smaller number of batches; hence, fewer optimization problems need to be solved.
Second, in many solution methods, warm-starting can be used to quickly solve the sampling approximations with overlapping samples, considerably reducing solution time.
In this paper, we consider a parametrized degree of overlap, where the degree of overlap is at the user's discretion, and examine this both theoretically and empirically. 
% THIS LAST SENTENCE NEEDS TO CHANGE!

@@

The amount of asymptotic reduction in $\var{\vartg{\ybb}}$ for a given $0\leq \gamma< m$ relative its nonoverlapping counterpart 
%$\var{\vtm{\ybb}}$ 
depends on the asymptotic ratio of $\gamma/m$, which we denote by $\gammab$. 
The maximally-overlapping case ($\gamma = 1$) corresponds to $\gammab=0$, and the nonoverlapping case corresponds to $\gammab=1$. 
\cite{damerdji1994strong,damerdji1995mean} studied batching structures that guarantee almost sure and mean-square consistency of $\vartg{\ybb}$ in the nonoverlapping and maximally-overlapping cases. 
Under appropriate conditions, these results, along with those of \cite{Meketon1984} and \cite{Welch1987}, establish that 
\begin{equation}
\label{eq:varvarred}
\frac{n}{m}\var{n \vartg{\ybb}} \rightarrow \beta_{\gammab} \sigma^4
\end{equation}
as the batch size $m$ and the number of batches $n/m$ tend to infinity. 
In (\ref{eq:varvarred}), $\beta_1 = 2$ for nonoverlapping batches, $\beta_0 = \frac{4}{3}$ for maximally overlapping batches, and more generally, $\beta_{\gammab}$ is twice the value of 
\begin{equation} \label{eq:var_reduct_formula}
	\gammab \left( 1 + 2 \sum_{j=1}^{\cogb-1} (1-j\gammab)^2 \right),
\end{equation}
where $\lceil \cdot \rceil$ denotes the smallest integer that is greater than or equal to its argument.
If $\ogb$ is an integer, (\ref{eq:var_reduct_formula}) simplifies to $\frac{2+\gammab^2}{3}$.
This means that the maximally-overlapping variance estimator has a variance that is two-thirds ($66.67\%$) of the nonoverlapping estimator.
As the amount of overlap decreases, the variance increases. 
When only half of the observations overlap ($\gammab = 1/2$ or $\gamma=m/4$), for instance, the variance is 75\% of the nonoverlapping counterpart.

@@

The optimality gap estimator for each batch $\gb_{lj}$ is defined like (\ref{eq:mrp_gap}) for general values of the nonoverlap parameter $\gamma_l$.  
We just removed the minimization in (\ref{eq:mrp_gap}) and used directly an optimal solution of $\xs_j$ of batch $j$.  
The overall mean, $\gbb_l$, is defined a little differently.  
Here, $\gbb_l$ still uses each observation $i = 1, 2, \dots, n_l$ but also makes use of {\it all} the information collected throughout the batches.  
That is, if observation $\xiti$ is used in $|B(i)|$ number of batches, all the optimal solutions $\xs_j$ corresponding to each batch $j \in B(i)$ are used for the lower bound estimator.  
Then, $VG_l$ is defined in a similar fashion for the overlapping batches variance estimator (\ref{eq:variance_gamma}).

@@

The internal optimization in the batches in (\ref{eq:gbar}) makes a straightforward statistical analysis of the behavior of the estimators difficult.  
To overcome this problem, we introduce the following unbiased optimality gap estimators:
\begin{align}
	\db_{lj} & = \frac{1}{m_l} \sum_{i=1}^{m_l} \left[ f(\xh,\xit^{\gamma_l(j-1)+i}) - f(\xs,\xit^{\gamma_l(j-1)+i}) \right],\ \ \ j = 1, 2, \dots, \nbl, \label{eq:dbar} \\
	\dbb_l & = \frac{1}{n_l} \sum_{i=1}^{n_l} \left[ f(\xh,\xiti) - f(\xs,\xiti) \right], \label{eq:dbb} \\
	VD_l & = \frac{1}{\left( \tfrac{n_l}{m_l} - 1 \right) \left(\nbl\right)} \sum_{j=1}^{\nbl} (\db_{lj} - \dbb_l)^2. \label{eq:vd}
\end{align}
These are essentially the same as the overlapping batches estimators in \S \ref{ssec:overlap} with $y^i = f(\xh,\xiti) - f(\xs,\xiti)$.  
These are also defined identically to the original estimators (\ref{eq:gbar})--(\ref{eq:vg}) with the exception that $\xs_j$ from (\ref{eq:gbar}) and (\ref{eq:gbb}) is replaced by an optimal solution $\xs$ in (\ref{eq:dbar}) and (\ref{eq:dbb}).  
With $\xh$ and $\xs$ fixed, estimators (\ref{eq:dbar})--(\ref{eq:vd}) have the same statistical properties as overlapping batches estimators in \S \ref{ssec:overlap}.  
The optimal solution $\xs$ is not known; however, the estimators (\ref{eq:dbar})--(\ref{eq:vd}) are used only to show convergence of (\ref{eq:gbar})--(\ref{eq:vg}) and not necessary for carrying out the OMRP algorithm.  
Throughout the rest of this section, we assume that the optimality gap estimators (\ref{eq:gbar})--(\ref{eq:vg}) and their nonoptimized counterparts (\ref{eq:dbar})--(\ref{eq:vd}) use identical samples and values of $n_
l$, $m_l$, and $\gamma_l$.  \smallskip
